\documentclass[a4paper]{report}
\input{notetemplate.tex}
\author{Christopher Fu}
\title{Econometrics I}

\thispagestyle{empty}
\addbibresource{ref.bib}

\usepackage{adjustbox}
\usepackage{centernot}
\usepackage{csquotes}

\begin{document}

\maketitle

\begin{abstract}
    This is a note template, with a simplified structure. Feel free to adjust for your usage.

    Now let's start a simple demo for you to take fancy notes in \LaTeX!
\end{abstract}

\newpage

\tableofcontents

\chapter{First Chapter}
% Your content here
\section{Useful Environment}
We now see some common environment you'll need to complete your note.

\begin{definition}[Natural number]\label{def}
  We denote the set of \emph{natural numbers} as \(\mathbb{\MakeUppercase{n}} \).
\end{definition}


\begin{lemma}[Useful lemma]\label{lma}
  Given the axioms of \hyperref[def]{natural numbers \(\mathbb{N}\)}, we have
  \[
    0\neq 1.
  \]
\end{lemma}
\begin{proof}[An obvious proof]
  Obvious.
\end{proof}
\begin{proposition}[Useful proposition]\label{prop}
  From \autoref{lma}, we have
  \[
    0<1.
  \]
\end{proposition}
\begin{exercise}
  Prove that \(1 < 2\).
\end{exercise}
\begin{answer}
  We note the following.
  \begin{note}
    We have \autoref{prop}! We can use it iteratively!
  \end{note}
  With the help of \autoref{lma}, this holds trivially.
\end{answer}
\begin{eg}
  We now can have \(a < b\) for \(a < b\)!
\end{eg}
\begin{explanation}
  Iteratively apply the exercise we did above.
\end{explanation}
\begin{remark}
  We see that \autoref{prop} is really powerful. We now give an immediate application of it.
\end{remark}

\begin{theorem}[Mass-energy equivalence]\label{thm}
  Given \autoref{prop}, we then have
  \[
    E = mc^2.
  \]
\end{theorem}
\begin{proof}
  The blank left for me is too small,\footnote{\url{https://en.wikipedia.org/wiki/Richard_Feynman}} hence we put the proof in \autoref{appendix}.
\end{proof}

\chapter{Second Chapter}
% Your content here

\newpage
%─────Appendix────────────────────────────────────────────────────────────────────────────────────────────────────────────────────────────────────────
\appendix
\appendixpage

\newpage
%─────Reference──────────────────────────────────────────────────────────────────────────────────────────────────────────────────────────────────────
\printbibliography

\end{document}
