\section{Neoclassical Growth Model}

The same basic environment as Solow model without the assumption of the constant exogenous saving rate.
\[ Y = F(K(t), L(t), Z(t)) \]

As in basic general equilibrium theory, let us suppose that preference orderings can be represented by utility functions. In particular, suppose that there is a unique consumption good, and each household $h$ has an \textit{\textcolor{blue}{instantaneous utility function}} given by: $u(c^h(t))$, where $c^h(t)$ is the consumption and $u : \mathbb{R}_+ \rightarrow \mathbb{R}$ is increasing and concave.

\begin{note}
\textcolor{red}{Negative levels of consumption are not allowed.}
\end{note}

We shall make 2 major assumptions:

\begin{description}[nosep]
\item[1.] Household does not derive any utility from the consumption of other households, so consumption externalities are ruled out.
\item[2.] Impose the condition that overall utility is \textcolor{blue}{\textit{time-separable and stationary}}; that is, instantaneous utility at time $t$ is independent of the consumption levels at past or future dates and is represented by the same utility function $u^h$ at all dates.
\end{description}

Representative household with preferences:
\[ U^h(T) = \sum_{t = 0}^T (\beta^h)^t u^h(c^h(t)) \]
where $\beta^h \in (0, 1)$ is the time discount factor of household $h$.

A solution $\{x(t)\}_{t = 1}^T$ to a dynamic optimization problem is \textit{\textcolor{blue}{time-consistent}} if the following is true: \textit{When $\{x(t)\}_{t = 1}^T$ is a solution to the continuation dynamic optimization problem starting from $t = 0$, $\{x(t)\}_{t = t'}^T$ is a solution to the continuation dynamic optimization starting from the time $t = t' > 0$}.

Let's consider the simplest case and suppose all households are infinitely-lived and identical. Then the demand side of the economy can be represented as the solution of the following maximization problem at time $t = 0$:
\[ \max \sum_{t = 0}^{\infty} \beta^t u(c(t)) \]
where $\beta \in (0, 1)$.

Budget constraint:
\[ C(t) + K(t + 1) = (1 + r(t))K(t) + w(t)L(t) \]
$r(t)$ is the rate of return on lending capital to firms.

\begin{note}
\textbf{\textcolor{blue}{Optimal Growth}}

\textbf{If the economy consists of a number of identical households, then this problem corresponds to the Pareto optimal allocation giving the same (Pareto) weight to all households. Therefore the optimal growth problem in discrete time with no uncertainty, no population growth, and no technological progress can be written as follows:}
\begin{align*}
    \max_{\{c(t), k(t)\}_{t=1}^{\infty}} &\sum_{t = 0}^{\infty} \beta^t u(c(t)) \\
    \text{s.t.} \quad &k(t + 1) = f(k(t)) + (1 - \delta)k(t) - c(t)
\end{align*}
with $k(t) \geq 0$ and given $k(0) > 0$.

The constraint is straightforward to understand: \textit{\textcolor{blue}{total output per capita produced with capital-labor ratio $k(t)$, $f(k(t))$, together with a fraction $(1 - \delta)$ of the capital that is undepreciated make up the total resources of the economy at date $t$.}}

Assuming that the representative household has $L(t)$ unit of labor supplied inelastically and denoting its assets at time $t$ by $a(t)$, this problem can be written as:
\begin{align*}
    \max_{\{c(t), k(t)\}_{t=1}^{\infty}} &\sum_{t = 0}^{\infty} \beta^t u(c(t)) \\
    \text{s.t.} \quad &a(t + 1) = (1 + r(t))a(t) - c(t) + w(t)L(t)
\end{align*}
where $r(t)$ is the net rate of return on assets, so that $1 + r(t)$ is the gross rate of return, and $w(t)$ is the equilibrium wage rate. \textcolor{red}{Market clearing then requires $a(t) = k(t)$.}
\end{note}

And, we have the non-Ponzi condition:
\[ \lim_{t \to \infty} \left\{K(t)\left[\prod_{s = 1}^{t-1} \left(\frac{1}{1 + r_s}\right)\right]\right\} \geq 0 \]
which refers to: limit of the PDV of debt has to be nonnegative (you cannot die in debt)

\begin{theorem}[\textcolor{blue}{\textbf{Euler Equation}}]
\[ \forall t : u'(c_t) = \beta(1 + r_{t + 1})u'(c_{t + 1}) \]
\end{theorem}

With an infinite horizon ($T \to \infty$), the Euler equations still determine the relative consumption levels. The terminal condition becomes \textcolor{blue}{transversality condition} is the limit of the terminal condition:
\[ \lim_{T \to \infty} K_{T + 1}\beta^T u'(c_T) = \lim_{T \to \infty} K_{T + 1}\beta^{T + 1}u'(c_{T + 1})(1 + r_{T + 1}) = 0 \]
Hence
\[ \lim_{t \to \infty} \beta^t u'(c_t)(1 + r_t)K_t = 0 \]
Then, we have two conditions involve the infinity:
\begin{equation}
    \lim_{t \to \infty} \beta^t u'(c_t)(1 + r_t)K_t = 0 \label{eq:trans}
\end{equation}
\begin{equation}
    \lim_{t \to \infty} \left\{K(t)\left[\prod_{s = 1}^{t-1} \left(\frac{1}{1 + r_s}\right)\right]\right\} \geq 0 \label{eq:nonponzi}
\end{equation}
The \eqref{eq:trans} is a condition for optimal behavior. The \eqref{eq:nonponzi} is a condition to make sure that the flow budget constraints are consistent with a lifetime budget constraint and no perpetual debt.

\textbf{Steady State} $(C^*, K^*)$

From the Euler equation:
\[ u'(C^*) = \beta(1 + F_K(K^*, L))u'(C^*) \]
hence
\[ 1 + r^* = 1 + F_K(k^*, 1) - \delta = \frac{1}{\beta} \]
which fully determines $k^*$.

\begin{note}
This shows that $k^*$ does not depend on $u(\cdot)$.
\end{note}

Level of per-capita consumption is then given by the economy's resource constraint:
\[ C^* + K^* = (1 - \delta)K^* + F(K^*, L) \]
hence,
\[ c^* = F(k^*, 1) - \delta k^* \]
\[ F_K(k^*, 1) = \delta + \left(\frac{1}{\beta} - 1\right) \]
In the Solow model the golden-rule level of capital (which maximizes steady-state consumption) satisfies
\[ F_K(k_S^{\text{GR}}, 1) = \delta \]
hence
\[ k^*_{\text{NGM}} < k_S^{\text{GR}} \]

\begin{note}
\textbf{Intuition:}

In NGM model, people tend to consume more on the near future rather than in the far steady state.

The simple reason is discounting: in the Ramsey model the planner puts less weight on the long-run
(anticipating that accumulating a lot of capital will require consumption sacrifices in the transition), 
while under the Golden Rule only the long-run outcome counts.
\end{note}

\section{NGM in continuous time}

Assume labor supply $L$ growth at rate $n$:
\[ L(t) = e^{nt} \]
Households choose consumption/saving to maximize
\[ U = \int_0^{\infty} e^{-\rho t}e^{nt}u(c(t))dt = \int_0^{\infty} e^{(n-\rho)t}u(c(t))dt \]
Intertemporal budget constraint: as in the discrete time case, but treat $\Delta K \to 0$:
\[ \dot{K}(t) = w(t)L(t) + (1 + r(t))K(t) - C(t) \]
Write in per-capita units:
\[ \dot{k}(t) = w(t) + (1 + r(t))k(t) - c(t) - nk(t) \]
Non-Ponzi condition:
\[ \lim_{t \to \infty} \left\{k(t)\exp\left[-\int_0^t(r(v) - n)dv\right]\right\} \geq 0 \]
To solve this, we define the \textbf{Hamiltonian}:
\[ H(c, k, \mu) = e^{-(\rho - n)t}u(c(t)) + \mu(t)[w(t) + (1 + r(t) - n)k(t) - c(t)] \]
optimality conditions are
\begin{align*}
    H_c(c, k, \mu) &= 0 \\
    H_k(c, k, \mu) &= -\dot{\mu}(t) \\
    H_{\mu}(c, k, \mu) &= \dot{k}(t) \\
    \lim_{t \to \infty} \mu(t)k(t) &= 0
\end{align*}
Hence:
\begin{align*}
    e^{-(\rho - n)t}u'(c(t)) - \mu(t) &= 0 \\
    \mu(t)(1 + r(t) - n) &= -\dot{\mu}(t)
\end{align*}
and we get:
\[ e^{-(\rho - n)t}u'(c(t))(1 + r(t) - n) = -e^{-(\rho - n)t}u''(c(t))\dot{c}(t) + (\rho - n)e^{-(\rho - n)t}u'(c(t)) \]
Denote $\sigma(c) = -\frac{u'(c)}{u''(c)c}$, we have the Euler equation in continuous time:
\[ \sigma(c)(1 + r(t) - \rho) = \frac{\dot{c}(t)}{c(t)} \]
$\sigma(c)$ is the intertemporal elasticity of substitution (IES).

\section{From NGM to Endogenous growth}

Key challenge: if capital is accumulated, and it has decreasing
returns.

Then how can we keep marginal product high?

\subsection{The AK model based on Arrow (1962) and Frankel (1962)}
Arrow’s (1962) approach to endogenous growth was to propose that technological progress is an un-
intended consequence of producing new capital goods, a phenomenon dubbed “learning by doing”.

\subsection{Romer (1980): increasing returns are a result of ideas being non-rival}
\begin{itemize}
    \item rival good: my consumption precludes you from consuming it
    \item non-rival good: all of us con consume it simultaneously without
    affecting others
\end{itemize}

Rivalry is a technological characteristic, whereas excludability is an
institutional characteristic: Non-excludable good: impossible to prevent it from being
consumed/used by other agents.

Non-rivalry naturally suggests increasing returns to scale.
\begin{eg}
    \

    Cost of producing a movie is high and a one-time investment. \\
    But once you produced it: constant returns to scale in DVD production,\\
    There's IRS in “movies on DVD”,\\
    Movie itself is non-rivalrous!\\
    Online piracy suggests that it may even not be (perfectly) excludable.
\end{eg}
Hence, the natural structure of endogenous growth is:
\[Y_t = F(A_t, K_t, L_t)\]
with
\begin{align*}
    F(A_t, \lambda K_t, \lambda L_t) &= \lambda F(A_t, K_t, L_t) \\
    F(\lambda A_t, \lambda K_t, \lambda L_t) &> \lambda F(A_t, K_t, L_t)
\end{align*}
Constant returns in rivalrous factors, increasing returns in all factors.

NOw, let's take the economy as a whole in a perfect competition setup with the
previous production function. Then the profit in the economy is zero:
\[\Pi _t = F(A_t, K_t, L_T) - w_t L_t - R_t K_t = 0 \]
as in Solow model/Ramsey model.

So how do we pay for the fixed cost of investment in ideas? Who
pays the research labs?

There are two solutions:

\begin{enumerate}
    \item Ideas accumulate as a by-product of production $\rightarrow$ nobody has to pay
    (Romer 1986, Lucas 1988, learning-by-doing)
    \item Give producers some market power $\rightarrow$ positive ex-post profits $\rightarrow$ can
    pay for research (Romer, 1990)
\end{enumerate}

\subsection{Models of R\&D}
Most simple models of endogenous technological change are those in
which R\&D expands the variety of inputs or machines used in
production (e.g. Romer 1990) due to the increasing number of varieties,
the division of labor(specialization) increases.

There's an alternative explanation: R\&D improves the quality of inputs or machines used in
production.

\underline{product innovation: }increases the number of inputs or the quality
of inputs.

\underline{process innovation: }reduces the cost of existing products.

\subsection{Sketch of Romer (1990) expanding varieties model}

Romer assumed there are three decision makers in the model:
\begin{enumerate}
    \item Households: maximize utility, subject to the usual budget constraint;
    \item Final good producers: hire labor and purchase intermediate input,
    and combine them to produce the final output, sold at unit price
    \item R\&D firms:
        \begin{itemize}
            \item Devote resources to invent new intermediate inputs
            \item Once a product has been invented, the innovating firm obtains a
            perpetual patent(allows the firm to sell at the price it chooses)
            \item price is chosen to maximize profits
        \end{itemize}
\end{enumerate}

So, the development intuition in this model is:

Higher willingness to save (lower rate of discounting $\rho $ or higher
elasticity of intertemporal substitution) leads to \underline{higher supply of capital},
so \underline{more resources invested into R\&D}, resulting in
\underline{higher rate of technological progress}, thus, there will be higher growth rate in the steady state.

The fact that the new “idea” can be used in a nonrival manner is key
for the R\&D output affecting productivity (or more generally, the cost
of production).

\subsection{Schumpeterian endogenous growth models}
Subsequent work used this mechanism to talk about the “industrial
organization” of the model (Aghion-Howitt 1992)

\begin{itemize}
    \item Stronger IP protection incentivizes firms to do R\&D;
    \item But IP protection also causes a markup distortion: patents give
    market power
    \item And innovation potentially displaces an incumbent (“creative
    destruction”), which again somewhat lowers the returns to R\&D.
    \item If factor reallocation is not working well, this can have additional
    costs.
\end{itemize}

\subsection{Problems with endogenous growth models}

At the heart of endogenous growth theories are mechanisms that are
incredibly hard to observe and measure:

In the case of Romer it all relies on the R\&D production function
(and in Aghion-Howitt also on market structure/competition). But
R\&D is a process where we observe neither inputs nor outputs well.
\begin{enumerate}
    \item Little of research output gets patented, and much of what gets patented is not economically valuable
    \item Difficult to distinguish R\&D input from production inputs with available data
    \item Market structure/conduct: even more difficult to observe (see:
    antitrust/IO).
\end{enumerate}

In the end, EGT use metaphors (“production of ideas”) that render the theory not falsifiable. 
And if it’s not falsifiable, it’s not scientific (Popper, 1934).